---
title: "Lab6WithWork"
author: "Rachel Behm"
date: "February 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

goals:
-read in spatial data
-simplify polygons so doesnt take 9 years to load
-add and transform projections
-make maps
-join spatial data
-find intersections
-interactive plots
-create sf from lat/lon

```{r}
library(tidyverse)
library(tmap)
library(sf)
library(leaflet)
library(ggrepel)
library(raster)
library(ggspatial)
library(RColorBrewer)
```

File types:
dbf attributes stored
prj projection stored
shp shapefile

###Example 1. Dams in California
```{r}
#simplify before running code or else sooooo slow

#working director(since in a project "." is all we need + prefix (read_sf)
#select takes attribute AND geometry for you woo thank you based sticky geometry
#simplify polygon so dont have to wait for map- collape points to single line if very close- makes much less points (st_simplify)
#set coordinate reference system to 4326 which is same as WGS84 (st_transform)

ca_eco <- read_sf(".",layer="ca_eco") %>%
  dplyr::select(US_L3NAME) %>%
  rename(Region = US_L3NAME) %>%
  st_simplify(dTolerance = 100) %>%
  st_transform(crs = 4326)


```

```{r}
#wont originally show an st_crs since didnt come with one, we have to set it
ca_counties <- read_sf(".", layer = "california_county_shape_file")

st_crs(ca_counties) = 4326
```

```{r}
ca_dams <- read_sf(dsn = ".", layer = "California_Jurisdictional_Dams") %>% # Read data
  rename(Condition = Condition_) # Change column name (remove final _)

ca_dams$Condition <- fct_relevel(ca_dams$Condition, "Fair","Satisfactory","Unsatisfactory","Poor") # Set factor levels (not sure if using this later...)
```

MAPSSSS
```{r}
plot(ca_eco)
#fast cause only use 1 attribute (name)
plot(ca_counties)
#plot in base R takes longer cause plot all attributes separate
```

now lets use ggplot
```{r}
#Rcolorbrewer doesnt have enough colors by default (only 10) we have to set it (since we have 13 regions)
color_count <- 13
#usecolorramppalette to make own palette (CRP overrides default number of colors)
my_colors <- colorRampPalette(brewer.pal(10, "Set2"))(color_count)


#color ="NA" makes the borders go away
#second geom_sf to get names from diff file
#they are perfectly overlayed since using same coordinate system
#make points that match these long and lat(our dam locations)
#can make raticles goaway with datum= NA

ggplot(ca_eco) +
  geom_sf(aes(fill = Region),
          color = "NA",
          show.legend = FALSE) +
  scale_fill_manual(values = my_colors)+
  geom_sf(data = ca_counties,
          fill="NA",
          color= "gray30",
          size=0.1) +
  geom_point(data= ca_dams, 
             aes(x=Longitude, y= Latitude),
             size=1,
             color="gray10",
             alpha=0.5)+
  theme_minimal() +
  coord_sf(datum = NA)


```

### Example 2. Damns in the Sierra Nevada Ecoregion
```{r}
#now join ecoregion and dam info (filter ecoregion)

sn <- ca_eco %>% 
  filter(Region == "Sierra Nevada") %>% 
  st_join(ca_dams)

# Then plot:
ggplot(sn) +
  geom_sf(data = ca_counties, fill = "wheat3", color = "NA") +
  geom_sf(fill = "lemonchiffon4", color = "NA") +
  geom_point(aes(x = Longitude, y = Latitude), size = 0.5, color = "red4") +
  theme_void() +
  coord_sf(datum=NA) +
  labs(x = "", y = "", title = "CA Dams in Sierra Nevada Eco-Region")
```

####Example 3. Santa Barbara County eco-regions

Can plot just pieces using st_intersection (for example, if we only want to plot eco-regions in Santa Barbara County), and crop graphing space with coord_sf() limits. 

```{r}

# Get just SB county

sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara")
#we are weird cause we have islands!

#now lets make clipped version that only shows intersection with SB county
eco_clip <- st_intersection(ca_eco, sb)

plot(eco_clip)
#just show Santa Barbara county!!!! super cool


#we still use the full ca_counties because we want them shown but just without ecoregions
#dont have to specify second geom_sf since use same clip
#its too big and we want a smaller area -> limit coordinates of map (bounding box)
#now it is cropped to a better looking area
#give context by manually crating a point for downtown SB
#we know limits and legend position just by trial and error of what looks good
ggplot(eco_clip) +
  geom_sf(data = ca_counties, fill = "gray90", color = "gray80", size= 0.2) +
  geom_sf(aes(fill = Region), color ="NA") +
  scale_fill_manual(values = c("darkolivegreen2", "darkolivegreen", "gold2")) +
  coord_sf(xlim= c(-121,-119), ylim = c(33.5,35.5)) +
  geom_point(aes(x= -119.6982, y= 34.4208)) +
  geom_text(x= -119.6982, y= 34.35, label = "Santa Barbara") +
  theme_minimal() +
  theme(legend.position = c(0.5,0.15))

#x and y labels look weird so would have to rename with labs
```

###Example 4. Interactive plots with tmap
```{r}
#make tmap object (tm_shape) then plot interactively
#syntax a lil diff then ggplot

map_sb_eco <- tm_shape(eco_clip) +
  tm_fill("Region", 
          palette= "RdPu", 
          alpha = 0.5) +
  tm_shape(ca_counties) +
  tm_borders() 


#makes it interactive
#also maintained in knitted doc
tmap_mode("view")

map_sb_eco


#can change basemap via leaflet::providers 
tm_basemap("CartoDB.DarkMatter") +
  tm_shape(eco_clip) +
  tm_borders(col= "white")
```

### Example 5. Fault Lines
```{r}

fault_lines <- read_sf(".", layer = "GMC_str_arc") %>%
  st_transform(crs = 4326) %>%
  separate(LTYPE, into = c("syn_ant", "certainty", "direction"), sep = ",")

plot(fault_lines)


ggplot() +
  geom_sf(data=ca_counties, fill = "black", color = "NA") +
  geom_sf(data = fault_lines, aes(color = syn_ant)) +
  theme_void()


#now if we just wanna see sb county fault lines
#diff way then precious
sb_faults <- fault_lines %>%
  st_intersection(sb)
ggplot()+
  geom_sf(data= sb) +
  geom_sf(data = sb_faults, aes(color = syn_ant))

#now what if we want interactive?
#dont have to set to view again 
tm_basemap("CartoDB.DarkMatter") +
  tm_shape(sb) +
  tm_borders(col="gray50", lwd= 2) +
  tm_shape(sb_faults) +
  tm_lines(col = "syn_ant", palette = c("orange","purple"), lwd=2)
```


###EXAMPLE 6. faceted maps
for example by year
```{r}
ggplot()+
  geom_sf(data = ca_counties, fill="black", color = "NA") +
  geom_sf(data= fault_lines, aes(color= syn_ant)) +
  facet_wrap(~syn_ant)
```

###Example 7. create spatial info from lat lon (haha i do that)- MAKING SPATIAL POINTS

```{r}
#x  y and lat lon are the same in this case
ca_sites <- read_csv("cadfw_sensitive_sites.csv")

#makes coords into sticky geometry
sites_sf <- st_as_sf(ca_sites, coords = c("LONDD", "LATDD"), crs=4326)

#check to make sure it works
#can use geom_sf instead of geom_point since converted them
ggplot()+ 
  geom_sf(data= ca_counties, fill="gray40")+
  geom_sf( data = sites_sf, aes(color= PRI_CODE), size= 0.3)

```

###Example 6. choropleth of CA counties by # of dams in each
ex number of oph genera per county
```{r}
#line up spaces by joining them 
intersection <- st_intersection(ca_dams, ca_counties)

#now count up number in each county
#group by county nam then tally them (counts observations in each group)
dams_per_county <- intersection %>%
  group_by(NAME) %>%
  tally()
```

